Experience Report: Haskell in Computational Biology

### 2012-03-08 mrfy_experience_report.tex
flushing our section 6

\subsection{What you must know to succeed}
\remark{I don't like this. I really struggled with this section. ---AG}
If others would like to follow in our footsteps, there are a few
things we believe one must know in order to succeed.
Firstly, some experience with functional programming concepts is a necessity. 
Specifically, coming from an imperative background, writing algorithms and 
modeling data without the ability to mutate state were the most difficult 
concepts to adapt to.
NEED MORE.

\subsection{Evaluation}
\remark{I think we need more in this section. ---AG}
In hindsight, we feel that there are several key facets of our project in 
Haskell that deserve evaluation.
Firstly, we found that existing tools and libraries were plentiful and 
typically helped us greatly.
In particular, Data.Memocombinators made dynamic 
programming with Viterbi exquisitely simple and elegant.
Control.Parallel.Strategies also made the parallelization of Viterbi trivial.
Most of the tools we used came straight from Hackage, and although 
documentation of some form was almost always present, it was sometimes 
difficult to tell if a particular tool was right for us.
In particular, parts of the documentation were inaccessible to us as beginning 
functional programmers.
We found that documentation with straight-forward examples to be the most 
helpful.

Another component worthy of evaluation is the debugging process.
Debugging in Haskell often takes the form of fixing type errors; which we think 
is a great thing.
However, in the case where debugging extends into runtime behavior, there 
doesn't appear to be any clearly defined approach to effective debugging.
Stack traces are available via profiling, but this was difficult to find and 
as we mentioned, profiling has its own problems.

>>>>

\subsection{What you must know to succeed}
\remark{I don't like this. I really struggled with this section. ---AG}
If others would like to follow in our footsteps, there are a few
things we believe one must know in order to succeed.
Firstly, some experience with functional programming concepts is a necessity. 
Specifically, coming from an imperative background, writing algorithms and 
modeling data without the ability to mutate state were the most difficult 
concepts to adapt to.

Secondly, identifying existing tools and libraries from Hackage that are mature 
is also important.
We found that some packages that could have been immensely useful---like one 
that provides backtraces---were either not maintained or contained bugs.
Some other packages like Pads were functional but immature, which resulted in 
additional code to work around the parts of Pads that had not been implemented 
yet.

Finally, a ``local expert'' or someone available to talk to with functional 
programming expertise, can make the transition to a language like Haskell 
easier.
We don't believe a local expert is absolutely necessary, but answers 
specifically related to functional programming idioms were easier to find when 
a local expert was available.

\subsection{Evaluation}
\remark{I think we need more in this section. ---AG}
In hindsight, we feel that there are several key facets of our project in 
Haskell that deserve evaluation.
Firstly, we found that existing tools and libraries were plentiful and 
typically helped us greatly.
In particular, Data.Memocombinators made dynamic 
programming with Viterbi exquisitely simple and elegant.
Control.Parallel.Strategies also made the parallelization of Viterbi trivial.
Most of the tools we used came straight from Hackage, and although 
documentation of some form was almost always present, it was sometimes 
difficult to tell if a particular tool was right for us.
In particular, parts of the documentation were inaccessible to us as beginning 
functional programmers.
Documentation that was accessible to us typically contained straight-forward 
examples showing how to use that particular tool or library.

Another component worthy of evaluation is the debugging process.
Debugging in Haskell often takes the form of fixing type errors; which we think 
is a great thing.
However, in the case where debugging extends into runtime behavior, we couldn't 
find any clearly defined approach to effective debugging.
Stack traces are available via profiling, but this was difficult to find and 
as we mentioned, profiling had its own barriers.

Finally, we believe that we could have achieved a better division of modules in 
our project, but that conflicted with one our goals: to get publishable results
quickly. Despite being hasty programmers, we were still able to produce code 
that is easy to understand and easily extensible. Namely, in our experience, 
hasty Haskell code beats hasty code in a language like C or C++.

In summary, with Haskell, we didn't have to master a completely different 
design and development style but we did have to wrap our heads around 
functional programming.
